---
title: "Ch2 Hmwk"
author: "Mac Faldet"
date: "February 22, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# An Introduction to Statistical Learning

## Chapter 2 Questions

### Question 1

(a) The sample size `n` is extremely large, and the number of predictors `p` is small.

  - We would expect a inflexible method ot perform better here. With many observations of few variables, we are likely to encounter many outliers. Since we can't intelligently explain this variance we are better off developing a more rigid and certain model.

(b) The number o fpredictors `p` is extremely large, and the number of observations `n` is small.

  - We would expect a fexible model to perform better with this amount of information. Vice Versa to (a).

(c) The relationship between the predictors and response is highly non-linear.

  - We would expect a fexible method to perform better. Rigid models depend on linear relationships, a  flexible model has build in methods to intelligently serve the unique underlying relations in the data.

(d) The variance of the error terms, i.e. `o`^2 = Var(e), is extremely high.

  - We would expect a inflexible method to perform better in this scenario. There is a high amount of variance, and a flexible model is likely to be swayed heavily by the outlier noise.
  

### Question 2 

(a) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.

  - This is a regression problem, and we are most concerned about inference. `n` = 500   `p` = 4

(b) We are launching a new product and wish to know whether it will be a `success` or a `failure`. We collect data on 20 similar products that were previously launched. For each product we hace recorded whether it was a success or failure, price charged for the product, marketing budget, competition price, and ten other variables.

  - This is a classification problem, and we are most concerned about prediction. `n` = 20   `p` = 14

(c) We are interested in predicting the % change in the USD/Euro exchange rate in relation to the weekly changes in the world stock markets. Hence we collect weekly data for all of 2012. For each week we record the % change in the USD/Euro, the % change in the US market, the % change in the British market, and the % change in the German market.

  - This is a regression problem, and we are concerned with prediction. `n` = 52   `p` = 4


### Question 10

```{r a}
library(MASS)
Boston
?Boston
```

(a) The Boston dataset has 506 rows and 14 columns. Each row is a town within Boston suburbia, each column is a town housing value.

```{r b}
pairs(Boston)
```

(b) From the pairwise plots you can see:

`crim`  correlates with `age`, `dis`, `rad`, `tax`, and `ptratio`
`zn`    correlates with `indus`, `nox`, `age`, and `lstat`
`indus` correlates with `age`   and `dis`
`nox`   correlates with `age`   and `dis`
`dis`   correlates with `lstat` and
`lstat` correlates with `medv`

(c) Are any related to per captia crime

```{r c}
plot(Boston$age, Boston$crim)
# Older homes, more crime
plot(Boston$dis, Boston$crim)
# Closer to work-area, more crime
plot(Boston$rad, Boston$crim)
# Higher index of accessibility to radial highways, more crime
plot(Boston$tax, Boston$crim)
# Higher tax rate, more crime
plot(Boston$ptratio, Boston$crim)
# Higher pupil:teacher ratio, more crime
```

(d) Any high outliers of crim, tax, ptratio?

```{r d}
hist(Boston$crim[Boston$crim>1], breaks=25)
# most cities have low crime rates, but there is a long tail: 18 suburbs appear
# to have a crime rate > 20, reaching to above 80
hist(Boston$tax, breaks=25)
# there is a large divide between suburbs with low tax rates and a peak at 660-680
hist(Boston$ptratio, breaks=25)
# a skew towards high ratios, but no particularly high ratios
```

(e) How many suburbs in this dataset cound the Charles river?

```{r e}
dim(subset(Boston, chas == 1))
# 35 suburbs
```

(f) Median ptratio?

```{r f}
median(Boston$ptratio)
# 19.05
```

(g) Lowest median owner occupied home? Comment on predictors.

```{r g}
t(subset(Boston, medv == min(Boston$medv)))
summary(Boston)
# The values on average are slightly better than other towns. 
```

(h) How many average above 7 rooms per dwelling? 8? Comment.

```{r}
dim(subset(Boston, rm > 7))
# 64
dim(subset(Boston, rm > 8))
# 13
summary(subset(Boston, rm > 8))
summary(Boston)
# lower crime, lower lstat, higher medv, low range tax
```

