---
title: "Intro - Iris"
author: "Mac Faldet"
date: "January 28, 2019"
output: html_document
---

# Welcome
## This is an Introduction to R and R Markdown

In this example we will be using the Iris dataset which is already downloaded to your system. 

Code chunks are contained within three consecutive tilde's. 
Within the brackets is:
    1st - type of code chunk
    2nd - name of this chunk
    3rd - unique settings for this code chunk environment
    
Name your code chunks so that you may reference them later and have all the settings carry over.

```{r customSetup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Above is the default preference when you open a new Rmarkdown document. 
To check out other customizable preferences for code chunks, run the following in an actual code chunk.   

"``{r options}
knitr::opts_chunk$get()
"``

Here's few good functions to run to understand your data better:

```{r chunckName, echo=FALSE}
#call the statistics summary of Iris
summary(iris)

#view the first 6 rows of Iris
head(iris)

#view the last 10 rows of Iris
tail(iris, nrow = 10)
```

Lets look at some simple plots provided by functions given by base r.

```{r plots}
#Check out the help window in the bottom left for more about plots.
?plot

# plot iris data with titles
plot(x=iris$Sepal.Length, y=iris$Sepal.Width, main="Iris Species", xlab="Sepal Length", ylab="Sepal Width", groups=iris$Species)

plot(x=iris$Species, y=iris$Sepal.Length, type="h", main="Iris Species Count", ylab="Sepal Length")
```

you can also use the `par()` function to split the graphical window, and put multiple graphs side by side.
```{r}
# plot with reduced margin and character sizes
par(mfrow = c(2,1), mex = 0.8, cex = 0.9)
plot(x=iris$Petal.Length, y=iris$Sepal.Length, main="Length Comparison", xlab="Petal", ylab="Sepal")
plot(x=iris$Petal.Width, y=iris$Sepal.Width, main="Width Comparison", xlab="Petal", ylab="Sepal")
```

It is often good practice to include the stats package for easy plotting of aggregations or statistical summaries.
```{r stat}
require(stats) # for lowess, rpois, rnorm
plot(iris)
lines(lowess(iris))
plot(sin, -pi, 2*pi) # see ?plot.function
```

Notice the variables going down the diagnol. 
The lowess lines makes it easy to see correlations between variables.

Correlations are very important, however the best way to compare correlations between variables is by throwing your numeric data into a correlation plot `corr()` provided by the `corrplot` package.
```{r correlationLoad, echo=F}
#Install corrplot package:
#install.packages("corrplot")
library(corrplot)
```

```{r correlation}
#index your numeric variables in your data set to pull them out for visualization
numeric_index <- is.numeric(iris)
iris_numeric <- iris[numeric_index,]

#view the coefficient matrix of variable eigenvalues to view the variable's correlations
I <- cor(iris_numerics)
head(round(I,2))

#plot this matrix to get more intuition about the relations
corrplot(I, method = "circle")
```


## Advanced Plots with ggplot2

If you need to download a package at any point, include [ install.packages(`desired_package`) ] in a code chunk.

For instance, we will use the `ggplot2` package. So you may need to run this below:    install.packages(ggplot2)

```{r install}
library(ggplot2)
```

To get a better understanding of the different flower types, lets stratify and compare their characteristics

```{r gg_sepals}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_jitter(alpha = 0.5) +
  stat_smooth(method = "lm", show.legend = T, col = "black") +
  coord_equal()

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(. ~ Species) +
  stat_smooth(method = "lm", se = F, fullrange = T, inherit.aes = T, col = "black") +
  coord_equal()
```

```{r gg_petals}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_jitter(alpha = 0.5) +
  stat_smooth(method = "lm", show.legend = T, col = "black") +
  coord_equal()

#Look at coord_equal to see the actual slope, it just minimizes the height of the graph so I left it off
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(. ~ Species) +
  stat_smooth(method = "lm", se = F, fullrange = T, inherit.aes = T, col = "black")# +
#  coord_equal()
```






## Case Study to Normalize Data

```{r packages needed to continue, echo=F}
?normalize
library(purrr)
library(BBmisc)
library(tidyverse)
```

```{r normalized comparison}
iris_numerics <- select(iris, -Species)
iris_normalized <- sapply(iris_numerics, normalize, method = "standardize")
head(iris_normalized)
iris_norm <- merge(iris_normalized, iris$Species)
colnames(iris_norm) <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")
head(iris_norm)
```

```{r advanced ggplots}
ggplot(iris_norm, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_line(alpha = 0.5) +
  facet_wrap(. ~ Species) +
  stat_smooth(method = "lm", se = F, fullrange = T, inherit.aes = T, col = "black") +
  coord_equal()
```
